---
title: "Userguide for the `vimixr` package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vimixr_userguide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(vimixr)
```

Let's generate some toy data. Here the data contains N = 100 samples, D = 2 dimensions and K = 2 clusters.
```{r data}
X <- rbind(matrix(rnorm(100, m=0, sd=0.25), ncol=2),
            matrix(rnorm(100, m=3, sd=1.5), ncol=2))
```


# Fixed-diagonal variance

```{r fixeddiag, message=FALSE, fig.height=4, fig.width=7}
  res <- cvi_npmm(X, variational_params = 20, prior_shape_alpha = 0.001,
          prior_rate_alpha = 0.001, post_shape_alpha = 0.001,
          post_rate_alpha = 0.001, prior_mean_eta = matrix(0, 1, ncol(X)),
          post_mean_eta = matrix(0.001, 20, ncol(X)),
          log_prob_matrix = matrix(runif(2000, -1, -0.001), 100, 20), maxit = 1000,
          fixed_variance = TRUE, covariance_type = "diagonal",
          prior_precision_scalar_eta = 0.001,
          post_precision_scalar_eta = matrix(0.001, 20, 1),
          cov_data = diag(ncol(X)))
k_post[n,1] <- res$posterior$`Cluster number`  
ari[n,1] <- mclust::adjustedRandIndex(true_l, apply(res$posterior$`log Probability matrix`, 1, which.max))
summary(res)
  plot(res)
```

# Varied diagonal variance
```{r varieddiag, message=FALSE, fig.height=4, fig.width=7}
 res <- cvi_npmm(X, variational_params = 20, prior_shape_alpha = 0.001,
            prior_rate_alpha = 0.001, post_shape_alpha = 0.001,
            post_rate_alpha = 0.001, prior_mean_eta = matrix(0, 1, ncol(X)),
            post_mean_eta = matrix(0.001, 20, ncol(X)),
            log_prob_matrix = matrix(runif(20*nrow(X), -1, -0.001), nrow(X), 20), 
            maxit = 1000,
            covariance_type = "diagonal",fixed_variance = FALSE,
            prior_shape_scalar_cov = 0.001,
            prior_rate_scalar_cov = 0.001,
            post_shape_scalar_cov = 0.001,
            post_rate_scalar_cov = 0.001,
            prior_precision_scalar_eta = 0.001,
            post_precision_scalar_eta = matrix(0.001, 20, 1))
  summary(res)
  plot(res)
```

# Fixed-full Variance
```{r fixedfull, message=FALSE, fig.height=4, fig.width=7}
  res <- cvi_npmm(X, variational_params = 20, prior_shape_alpha = 0.001,
          prior_rate_alpha = 0.001, post_shape_alpha = 0.001,
          post_rate_alpha = 0.001, prior_mean_eta = matrix(0, 1, ncol(X)),
          post_mean_eta = matrix(0.001, 20, ncol(X)),
          log_prob_matrix = matrix(runif(20*nrow(X), -1, -0.001), nrow(X), 20), 
          maxit = 1000,
          covariance_type = "full",fixed_variance = TRUE,
          post_cov_eta = array(rep(diag(ncol(X)), 20), c(ncol(X), ncol(X), 20)),
          prior_cov_eta = 1000*diag(ncol(X)),
          cov_data = diag(ncol(X)))
  summary(res)
  plot(res)
```

#Full Variance IW distribution
```{r fulliw, message=FALSE, fig.height=4, fig.width=7}
 res <- cvi_npmm(X, variational_params = 20, prior_shape_alpha = 0.001,
          prior_rate_alpha = 0.001, post_shape_alpha = 0.001,
          post_rate_alpha = 0.001, prior_mean_eta = matrix(0, 1, ncol(X)),
          post_mean_eta = matrix(0.001, 20, ncol(X)),
          log_prob_matrix = matrix(runif(20*nrow(X), -1, -0.001), nrow(X), 20), 
          maxit = 1000,
          covariance_type = "full",fixed_variance = FALSE,
          cluster_specific_covariance = FALSE,
          variance_prior_type = "IW",
          prior_df_cov = ncol(X) + 2,
          prior_scale_cov = diag(ncol(X))*100,
          post_df_cov = ncol(X) + 2,
          post_scale_cov = diag(ncol(X)),
          post_cov_eta = array(rep(diag(ncol(X)), 20), c(ncol(X), ncol(X), 20)),
          prior_cov_eta = 1000*diag(ncol(X)))
  summary(res)
  plot(res)
```

# Full Variance with Cholesky decomposition
 
```{r fulldecomposed, message=FALSE, fig.height=4, fig.width=7}
  res <- cvi_npmm(X, variational_params = 20, prior_shape_alpha = 0.001,
           prior_rate_alpha = 0.001, post_shape_alpha = 0.001,
           post_rate_alpha = 0.001, prior_mean_eta = matrix(0, 1, ncol(X)),
           post_mean_eta = matrix(0.001, 20, ncol(X)),
           log_prob_matrix = matrix(runif(20*nrow(X), -1, -0.001), nrow(X), 20), 
           maxit = 1000,
           covariance_type = "full",fixed_variance = FALSE,
           cluster_specific_covariance = FALSE,
           variance_prior_type = "decomposed",
           prior_shape_diag_decomp = 0.001,
           prior_rate_diag_decomp = 0.001,
           prior_mean_offdiag_decomp = 0,
           prior_var_offdiag_decomp = 1,
           post_shape_diag_decomp = matrix(0.001, 1, ncol(X)),
           post_rate_diag_decomp = matrix(0.001, 1, ncol(X)),
           post_mean_offdiag_decomp = matrix(0, 1, 0.5*ncol(X)*(ncol(X)-1)),
           post_var_offdiag_decomp = matrix(0.001, 1, 0.5*ncol(X)*(ncol(X)-1)),
           post_cov_eta = array(rep(diag(ncol(X)), 20), c(ncol(X), ncol(X), 20)),
           prior_cov_eta = 1000*diag(ncol(X)))
  summary(res)
  plot(res)
```

# Cluster-specific diagonal with IW distribution
```{r csiw, message=FALSE, fig.height=4, fig.width=7}
 res <- cvi_npmm(X, variational_params = 20, prior_shape_alpha = 0.001,
           prior_rate_alpha = 0.001, post_shape_alpha = 0.001,
           post_rate_alpha = 0.001, prior_mean_eta = matrix(0, 1, ncol(X)),
           post_mean_eta = matrix(0.001, 20, ncol(X)),
           log_prob_matrix = matrix(runif(20*nrow(X), -1, -0.001), nrow(X), 20), 
           maxit = 1000,
           covariance_type = "full",fixed_variance = FALSE,
           cluster_specific_covariance = TRUE,
           variance_prior_type = "IW",
           prior_df_cs_cov = ncol(X) + 2,
           prior_scale_cs_cov = diag(ncol(X)),
           post_df_cs_cov = matrix(rep(ncol(X) + 2, 20), nrow = 1),
           post_scale_cs_cov = array(rep(diag(ncol(X)), 20), c(ncol(X), ncol(X), 20)),
           scaling_cov_eta = nrow(X))
  summary(res)
  plot(res)
```

# Cluster-specific sparse Variance
```{r cssparse, message=FALSE, fig.height=4, fig.width=7}
 res <- cvi_npmm(X, variational_params = 20, prior_shape_alpha = 0.001,
           prior_rate_alpha = 0.001, post_shape_alpha = 0.001,
           post_rate_alpha = 0.001, prior_mean_eta = matrix(0, 1, ncol(X)),
           post_mean_eta = matrix(0.001, 20, ncol(X)),
           log_prob_matrix = matrix(runif(20*nrow(X), -1, -0.001), nrow(X), 20), 
           maxit = 1000,
           covariance_type="full",fixed_variance=FALSE,
           cluster_specific_covariance = TRUE,
           variance_prior_type = "sparse",
           prior_shape_d_cs_cov = matrix(rep(100, 20), nrow = 1, ncol = 20),
           prior_rate_d_cs_cov = matrix(rep(100, 20*ncol(X)), 20, ncol(X)),
           prior_var_offd_cs_cov = 1000,
           post_shape_d_cs_cov = matrix(0.001, 1, 20),
           post_rate_d_cs_cov = matrix(0.001, 20, ncol(X)),
           post_var_offd_cs_cov = matrix(0.001, 20, 3),
           scaling_cov_eta = nrow(X))
  summary(res)
  plot(res)
```

# Cluster-specific variance with off-diagonal normal priors
```{r csoffd, message=FALSE, fig.height=4, fig.width=7}
 res <- cvi_npmm(X, variational_params = 20, prior_shape_alpha = 0.001,
           prior_rate_alpha = 0.001, post_shape_alpha = 0.001,
           post_rate_alpha = 0.001, prior_mean_eta = matrix(0, 1, ncol(X)),
           post_mean_eta = matrix(0.001, 20, ncol(X)),
           log_prob_matrix = matrix(runif(20*nrow(X), -1, -0.001), nrow(X), 20), 
           maxit = 1000,
           covariance_type="full",fixed_variance=FALSE,
           cluster_specific_covariance = TRUE,
           variance_prior_type = "off-diagonal normal",
           prior_shape_d_cs_cov = matrix(rep(100, 20), nrow = 1, ncol = 20),
           prior_rate_d_cs_cov = 100,
           prior_var_offd_cs_cov = 0.0001,
           post_shape_d_cs_cov = matrix(0.001, 1, 20),
           post_rate_d_cs_cov = matrix(0.001, 20, ncol(X)),
           post_mean_offd_cs_cov = array(rep(diag(ncol(X)), 20), c(ncol(X), ncol(X), 20)),
           scaling_cov_eta = nrow(X))
  summary(res)
  plot(res)

```

